---
title: "Female survival after a Cold Shock"
author: "Marine Caussignac"
date: "2026-06-18"
output: html_document: code_folding: hide 
---


## **1. Introduction and Objective**
The objective of this analysis is to evaluate the impact of 3  chromosomal inversions (*3RP*, *2Lt* and *3RK*) on the survival rate of adult *Drosophila* females following exposure to a cold shock. 

### Experimental Design Overview

14 to 15 days old flies, were sexed using gas (CO₂). From these flies, vials containing exactly 15 males and 15 females were made. The vials were then immersed in ice at 0°C for 18 hours. 

Vials were covered with a plastic sheet to prevent the plugs from getting wet. After the shock, flies were allowed a recovery time of 4 hours.  Surviving individuals were then counted and sexed.

### Dataset

The dataset used for this analysis is the [ACS_F.csv](Dataset/ACS_F.csv) file. Its structure is describe in the [Data Readme] document.


## **2. Setup and Packages Loading**

For data manipulation, statistcal analysis and visualization, we load a selection of multiple library packages :

- **Data manipulation:** `dplyr`, `stringr`
- **Statistics:** `car` (Assumptions), `rstatix` (Non-parametric post-hoc tests)
- **Visualization:** `ggplot2`, `patchwork`

A standardized graphical theme (`plot_theme`) and a custom color palette (`Colors`) are also set up.The Y-axis limits are strictly set between 0.2 (20%) and 1.0 (100%) to ensure visual consistency across all panels.

```{r setuo, message=FALSE, warning=FALSE}
library(dplyr) 
library(ggplot2) 
library(rstatix) 
library(car) 
library(stringr)  
library(patchwork) 

Colors = c(
  "STD" ="#00b4d8",
  "3RP"="#f25c54",
  "3RP_HET_1"="#ffbf69",
  "3RP_HET_2"="#ffdab9",
  "2Lt"="#52b788",
  "2Lt_HET_1"="#80ed99",
  "2Lt_HET_2"="#c7f9cc",
  "3RK"="#8367c7",
  "3RK_HET_1"="#e0aaff",
  "3RK_HET_2"="#ebd9fc"
)

plot_theme <- theme_bw()+
  theme(axis.line = element_line(color = "black", linewidth = 0.8),
        axis.title.x = element_text(face = "bold", size = 22),
        axis.title.y = element_text(face = "bold", size = 22),
        axis.text.x = element_text(face = "bold", size = 16),
        axis.text.y = element_text(face = "bold", size = 16),
        legend.title = element_text(face = "bold", size = 20),
        legend.text = element_text(face = "bold", size = 18),
        plot.tag = element_text(size = 20, face="bold")
  )

y_axis <- scale_y_continuous(
  limits = c(0.2, 1),
  breaks = c(0.2, 0.4, 0.6, 0.8,1)
)
```


## 3. Data Preparation

The [ACS_F.csv](Dataset/ACS_F.csv) dataset is imported and a new biological variable is calculated for each replicate :

- **Female Survival Rate (`F.Survival`):** Calculated as the ratio of surviving females (`ALIVE.F`) to the total initial number of females (`TOTAL.F`).

```{r}

```


## 4. Assumption Checks

A standard linear model (`lm`) is fitted to extract residuals and test the assumptions required for parametric statistics (AONVA):

-  **Normality:** Evaluated statistically using the Shapiro-Wilk test and visually assessed via a two-panel graphical output (Panel A: Histogram of residuals; Panel B: Q-Q Plot).
- **Homoscedasticity:** The equality of variances across karyotype groups is tested using Levene's Test.

```{r}

```
 
As both the Levene's Test and the Shapiro-Wilk test remains significant, the assumptions required for a parametric test are not met. Therefore, a non-parametric is performed.

## 5. Statistical Analysis (Non-Parametric)

As non-parametric test wt posth, we made :

- **A Kruskal-Wallis test** : allow to detect if at least one of the karyotype differ from the other
* **Post-hoc Test:** A global Dunn's test is executed to identify specific pairwise differences (using an adjusted p-value threshold of < 0.05).


### Contrast Tests

To provide a clearer biological interpretation, the dataset is subset into four distinct groups (3RP, 2Lt, 3RK, and Homozygotes) to directly compare variants against the STD control.
* An unadjusted Dunn's test is applied to each subgroup independently.
* A custom critical alpha threshold is mathematically calculated to correct for 18 simultaneous multiple comparisons.
* Pairwise comparisons falling below this strict threshold are flagged with an asterisk (`*`) in a newly generated `p.signif` column.

## 8. Visualization
The survival distributions are visualized using four boxplots.
* **Panel A:** 3RP family vs. STD
* **Panel B:** 2Lt family vs. STD
* **Panel C:** 3RK family vs. STD
* **Panel D:** Homozygotes vs. STD
* *Note: As no comparisons met the highly stringent corrected significance threshold, p-value annotations were omitted from the final graphs to maintain visual clarity.*

The four individual plots are assembled into a single cohesive publication-ready figure using the `patchwork` package. The individual panel tags (A, B, C, D) are explicitly assigned using `labs(tag=...)` to ensure robust formatting. The final composite image is exported as a high-resolution PNG (300 dpi) with optimized dimensions (45x25 cm).